Pollution, agricultural productivity, and development: Evidence from coal plants in India^[I thank Marc Bellemare and David Sungho Park for comments and suggestions that greatly improved the quality of the paper. The usual disclaimer applies. All scripts used in this paper are publicly available on 

### 2023-03-02 draft/draft.Rmd
scripts and draft

---
title: "Pollution, agricultural productivity, and development: Evidence from coal plants in India^[I thank Marc Bellemare and David Sungho Park for comments and suggestions that greatly improved the quality of the paper. The usual disclaimer applies. All scripts used in this paper are publicly available on my GitHub page: github.com/JoshMerfeld/pollution_development]"
author:
  - Joshua D. Merfeld^[KDI School of Public Policy and Management and IZA; merfeld@kdis.ac.kr]
date: "`r Sys.Date()`"
abstract: |
  | \noindent I show that emissions from a common energy source lead to decreased productivity and growth. The rapid increase in population in India from the 1990s through the 2010s led to the rapid construction of new coal power plants throughout the country. Using this roll-out combined with random changes in wind direction, I document negative effects of pollution from coal power plants on agricultural productivity throughout the country. Results suggest this decrease may be driven by a decline in labor allocation of the oldest half of my sample, who greatly decrease labor allocated to both farm and non-farm employment when exposed to pollution. These impacts spill over into aggregate economic growth, as well; nightlight growth is slower following years of higher exposure to pollution.
  |
  | \textbf{\textit{Keywords}}: pollution, productivity, agriculture, labor, India
  | \textbf{\textit{JEL Codes}}: H40, I15, J22, O13, Q52, Q53

@@

# Tables
plantresultstable <- readRDS("pollution_development/draft/tables/plantresultstable.rds")
pollutiontable <- readRDS("pollution_development/draft/tables/pollutiontable.rds")
yieldtable <- readRDS("pollution_development/draft/tables/yieldtable.rds")
yieldtablepm <- readRDS("pollution_development/draft/tables/yieldtablepm.rds")
yieldtabletwo <- readRDS("pollution_development/draft/tables/yieldtabletwo.rds")
yieldtablehet <- readRDS("pollution_development/draft/tables/yieldtablehet.rds")
yieldtablehet1st <- readRDS("pollution_development/draft/tables/yieldtablehet1st.rds")
yieldtablemonthly <- readRDS("pollution_development/draft/tables/yieldtablemonthly.rds")
pmeffects <- readRDS("pollution_development/draft/tables/yieldtablePM.rds")
labortable <- readRDS("pollution_development/draft/tables/labortable.rds")
labortablerural <- readRDS("pollution_development/draft/tables/labortablerural.rds")
labortableurban <- readRDS("pollution_development/draft/tables/labortableurban.rds")
labortableruralyoung <- readRDS("pollution_development/draft/tables/labortableruralyoung.rds")
labortableruralold <- readRDS("pollution_development/draft/tables/labortableruralold.rds")
wagestable <- readRDS("pollution_development/draft/tables/wagestable.rds")
labortablemonsoon <- readRDS("pollution_development/draft/tables/labortablemonsoon.rds")
labortablewinter <- readRDS("pollution_development/draft/tables/labortablewinter.rds")
labortableold <- readRDS("pollution_development/draft/tables/labortableold.rds")
labortableoldstate <- readRDS("pollution_development/draft/tables/labortableoldstate.rds")
ntltable <- readRDS("pollution_development/draft/tables/ntltable.rds")
yieldtableleads <- readRDS("pollution_development/draft/tables/yieldtableleads.rds")
yieldtablepostplant <- readRDS("pollution_development/draft/tables/yieldtablepostplant.rds")
laborplot <- readRDS("pollution_development/draft/tables/laborplot.rds")
monthplot <- readRDS("pollution_development/draft/tables/monthplot.rds")

@@

I study the effects of emissions from coal power plants on agricultural productivity, labor allocation, and economic growth in India. Previous work has suggested that there may be harmful direct effects of air pollution [@heck1982assessment; @marshall1997hidden] and water pollution [@reddy2006impact] on agricultural productivity, including work showing that gold mines -- which release large amounts of pollutants into the surrounding environment -- lead to lower agricultural productivity in Ghana [@aragon2016polluting]. These effects can work through the well known effects on human health and productivity, but also directly through land productivity. Black carbon in soot, for example, can directly affect solar radiation reaching plants on the surface [@ramanathan2008global; @burney2014recent]. Relative to previous work in economics, this paper tackles the broader case of coal plants, which are a leading source of power throughout the world. I show that air pollution from these plants can lead to lower agricultural productivity and even decreased economic growth.

@@

To identify the effects of air pollution on agricultural productivity, I instead combine the roll out of coal plants in India with plausibly exogenous changes in wind direction -- in a similar spirit to @deryugina2019mortality -- to measure the effects of pollution from coal plants on key development outcomes. I first plot the location of coal plants in India by year. Using a village-level shapefile, I identify all villages that are located within 30km of a coal plant at any time during the sample period, which comprises the 1990s to the 2010s, with differences depending on data availability. To measure exposure to emissions from coal plants, I calculate the direction from each coal plant to all villages within 30km. Using daily wind data and these spatial variables, I create an identifier separately for each of the one-hundred-thousand-plus villages for whether a given village is located downwind from a coal plant on any given day. I then aggregate this exposure variable based on the desired analysis; for agricultural productivity, for example, I create an exposure variable measuring the total number of exposure days in the five-month agricultural season, separately for the main monsoon season and the smaller winter season. Consistent with expectations, wind direction is a strong predictor of particulate matter at the monthly level. More days being located downwind from a coal power plant leads to higher average concentrations of particulate matter 2.5.

Using newly available estimates of India-wide agricultural (land) productivity, results show that the total number of days of being downwind from a coal power plant has a negative effect on overall agricultural productivity. Just a single additional day of being downwind from a coal power plant leads to a decrease in predicted yields of 0.3 percent. The within-village absolute deviation of wind exposure is around eight days, meaning that median changes in wind direction are responsible for fluctuations in agricultural productivity of around 2.4 percent, with changes of more than five percent from season to season being common. Using wind direction as an instrument for particulate matter, an increase of PM 2.5 by one microgram per cubic meter leads to an even larger decrease in agricultural productivity, at around 20 percent. This IV strategy is similar to that in @deryugina2019mortality, though they use general "high-pollution" wind directions while I concentrate only on wind directions from the location of a coal power plant. The IV estimates indicate that the effect of pollution is more negative than naïve OLS estimates indicate, which is consistent with the evidence on the construction of coal plants; higher levels of pollution are correlated with higher levels of agricultural productivity and economic growth, a relationship that the IV strategy adjusts downwards.

These effects are apparently concentrated in villages that do not experience the most extreme exposure to pollution; villages with higher maximum exposure experience effects only around 10 percent as large as villages with lower maximum exposure. Additional heterogeneity analysis shows that, despite the results with respect to the maximum, the effects are increasing in *mean* exposure. I also show that the effects are largest in villages with the highest initial agricultural productivity and that pollution and rainfall shocks are compounding; higher exposure along with less rainfall increases the negative effect of either shock individually. As weather shocks become more common across the globe, the rise of pollution in India may be particularly problematic for farmers.

Robustness checks show that this effect is concentrated in the years after the construction of the power plant, which helps decrease concerns around endogeneity since, conditional on the village fixed effect and weather (rainfall and temperature), wind direction across years and seasons is exogenous. Additionally, leads for wind direction do not predict current agricultural productivity, again supporting the identifying assumptions.

While I am not able to explicitly test effects on land productivity, I look at how adults allocate labor over the last seven days using the National Sample Survey and creating a district-level exposure measure based on the timing of the survey for each household. An additional day of wind over the previous week decreases overall labor allocation by `r paste0(substr(labortable[1,2],2,6))` days. However, the overall results show stark heterogeneity based on age; the older half of the sample greatly decreases labor allocation in response to pollution exposure while the younger half increases overall labor allocation. Older adults decrease labor allocation to both farm and non-farm labor, while younger adults increase labor allocated to farm labor. This is suggestive evidence of productivity effects being driven by the health and productivity effects on individuals, since older adults are more susceptible to pollution and they also greatly decrease non-farm employment, which presumably will no suffer any possible direct effects of pollution on productivity. The preponderance of results seems to indicate that effects on labor are the most likely mechanism.

The final set of results explores whether there are impacts of pollution on aggregate economic development. Using nightlights, I show that lagged wind -- i.e. pollution exposure in the prior year -- has a negative impact on economic development as proxied by nightlights. Going from the 25th percentile to the 75th percentile of wind exposure is synonymous with a decrease in the growth of nightlights within a village by around 0.7 percent. While this is a relatively small magnitude, it is worth noting that there are more than 100,000 villages in this sample. While the effect for a single village might be small, the overall, aggregate effect on growth is quite substantial. To put this number in perspective, @castello2018higher argue that a standard-deviation increase in the share of adults with a university education leads to 12 percent higher nightlights in India, while @storeygard2016farther shows that a four-fold increase in oil prices over six years leads to seven percent lower nightlights at a distance of around 500km from the nearest large port in sub-Saharan Africa. In other words, the 0.7 percent change in nightlights is quite large given this is annual variation.

@@

```{r plants, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Coal plants in India from 1990 to 2010", fig.footer = list("Note: The top figure shows the location of coal plants in 1990. The bottom figure shows the location of coal plants in 2010.")}

Figure \@ref(fig:plants) displays the location of coal plants in India for two specific years: 1990 and 2010. There is a clear increase in the prevalence of coal plants across the country over the two decades. Additionally, the overall capacity from coal plants increased from `r paste0(reportDecOne(sum(plants1990_capacity$capacity)))` gigawatts to `r paste0(reportDecOne(sum(plants2010_capacity$capacity)))` gigawatts, an increase of `r paste0(reportPercent((sum(plants2010_capacity$capacity) - sum(plants1990_capacity$capacity))/sum(plants1990_capacity$capacity)))` percent in just 20 years. This increase is driven both by the roll out of new plants as well as the construction of new units at existing plants, the latter of which is not shown in Figure \@ref(fig:plants).

@@

[^10]: I define "towards" as within five degrees to help take into account that I am using village centroids, using the x and y components of wind speed.

>>>>

---
title: "Pollution, agricultural productivity, and development: Evidence from coal plants in India^[I thank Marc Bellemare and David Sungho Park for comments and suggestions that greatly improved the quality of the paper. The usual disclaimer applies. All scripts used in this paper are publicly available on my GitHub page: github.com/JoshMerfeld/pollution_development]"
author:
  - Joshua D. Merfeld^[KDI School of Public Policy and Management and IZA; merfeld@kdis.ac.kr]
date: "`r Sys.Date()`"
abstract: |
  | \noindent Pollution is a growing problem in many developing countries. Using the rollout of coal plants to identify high-pollution locations in India and random variation induced by changes in wind direction, I show that pollution has large, negative impacts on agricultural productivity. Results indicate that some of this impact is driven by changes in labor allocation patterns in response to increased exposure to pollution. I also show that pollution from coal plants, specifically, has a larger effect on agricultural productivity than other types of pollution.
  |
  | \textbf{\textit{Keywords}}: pollution, productivity, agriculture, labor, India
  | \textbf{\textit{JEL Codes}}: H40, I15, J22, O13, Q52, Q53

@@

# Tables
plantresultstable <- readRDS("pollution_development/draft/tables/plantresultstable.rds")
pollutiontable <- readRDS("pollution_development/draft/tables/pollutiontable.rds")
yield1reducedform <- readRDS("pollution_development/draft/tables/yield1reducedform.rds")
yield2naive <- readRDS("pollution_development/draft/tables/yield2naive.rds")
yield3ivmain <- readRDS("pollution_development/draft/tables/yield3ivmain.rds")
yield4monthly <- readRDS("pollution_development/draft/tables/yield4monthly.rds")
yield5heterogeneity <- readRDS("pollution_development/draft/tables/yield5heterogeneity.rds")
yield6leads <- readRDS("pollution_development/draft/tables/yield6leads.rds")
yield7ivdiffindiff <- readRDS("pollution_development/draft/tables/yield7ivdiffindiff.rds")
labor1all <- readRDS("pollution_development/draft/tables/labor1all.rds")
labor2rural <- readRDS("pollution_development/draft/tables/labor2rural.rds")
labor3month <- readRDS("pollution_development/draft/tables/labor3month.rds")
laborplot <- readRDS("pollution_development/draft/tables/laborplot.rds")
monthplot <- readRDS("pollution_development/draft/tables/monthplot.rds")
opplot <- readRDS("pollution_development/draft/tables/opplot.rds")

@@

I study the effects of emissions from coal power plants on agricultural productivity in India. Previous work has suggested that there may be harmful direct effects of air pollution [@heck1982assessment; @marshall1997hidden] and water pollution [@reddy2006impact] on agricultural productivity, including work showing that gold mines -- which release large amounts of pollutants into the surrounding environment -- lead to lower agricultural productivity in Ghana [@aragon2016polluting]. These effects can work through the well known effects on human health and productivity, but also directly through land productivity. Black carbon in soot, for example, can directly affect solar radiation reaching plants on the surface [@ramanathan2008global; @burney2014recent]. Relative to previous work in economics, this paper tackles the broader case of coal plants, which are a leading source of power throughout the world. I show that air pollution from these plants can lead to lower agricultural productivity and even decreased economic growth.

@@

To identify the effects of air pollution on agricultural productivity, I instead combine the roll out of coal plants in India with plausibly exogenous changes in wind direction -- in a similar spirit to @deryugina2019mortality -- to measure the effects of pollution from coal plants on key development outcomes. I first plot the location of coal plants in India by year. Using a village-level shapefile, I identify all villages that are located within 30km of a coal plant at any time during the sample period, which comprises the 1990s to the 2010s, with differences depending on data availability.

I use the location of coal plants to identify high-pollution areas and calculate the direction from each coal plant to all villages within 30km. I include all areas, regardless of whether the coal plant has opened yet, to allow me to disaggregate the effects of pollution from coal plants with other kinds of pollution. Using daily wind data and these spatial variables, I create an identifier separately for each of the one-hundred-thousand-plus villages for whether a given village is located downwind from a coal plant on any given day. I then aggregate this exposure variable based on the desired analysis; for agricultural productivity, for example, I create an exposure variable measuring the total number of exposure days in the five-month agricultural season, focusing only on the monsoon (Kharif) season. Consistent with expectations, wind direction is a strong predictor of particulate matter at the monthly level. More days being located downwind from a coal power plant leads to higher average concentrations of particulate matter 2.5.

Using newly available estimates of India-wide agricultural (land) productivity, results show that the total number of days of being downwind from a coal power plant has a negative effect on overall agricultural productivity. Just a single additional day of being downwind from a coal power plant leads to a decrease in predicted yields of 0.6 percent. Combined with the within-village absolute deviation of wind exposure, results indicate that median changes in wind direction are responsible for fluctuations in agricultural productivity of around 1.5 percent.

Using wind direction as an instrument for particulate matter, an increase of PM 2.5 by one percent leads to an even larger decrease in agricultural productivity, at around 0.73 percent. Given within-village changes across years, pollution could be responsible for swings of 7-to-14 percent in agricultural productivity across years. This IV strategy is similar to that in @deryugina2019mortality, though they use general "high-pollution" wind directions while I concentrate only on wind directions from the location of a coal power plant. The IV estimates indicate that the effect of pollution is more negative than naïve OLS estimates indicate, which is consistent with the evidence on the construction of coal plants; higher levels of pollution are correlated with higher levels of agricultural productivity and economic growth, a relationship that the IV strategy adjusts downwards.

These effects are apparently larger in villages that experience higher median exposure to pollution and in villages with higher initial yields. I also show that rainfall shocks are compounding; higher exposure along with less rainfall increases the negative effect of either shock individually. As weather shocks become more common across the globe, the rise of pollution in India may be particularly problematic for farmers.

To explore the mechanisms driving this effect, I also use data on labor allocation from the National Sample Survey (NSS). While there do not appear to be large changes in labor allocation patterns when looking at the season as a whole, there is a clear pattern when we break down exposure to pollution by month: higher levels of pollution at the beginning of the monsoon season (June and July) lead to lower levels of farm labor allocation, but no changes in non-farm labor allocation. This matches up well with a similar exercise with agricultural yields: pollution in June is the only month in which there is a negative effect of pollution on agricultural yield. In tandem, these two results point to effects on labor as one mechanism driving the total effects, though I cannot rule out direct effects of pollution on land productivity.

The final set of results tries to disentangle the effects of pollution from the effects of coal pollution, specifically. Combining instrumental variables with a differences-in-differences strategy, I show that agricultural productivity has an elasticity of around -0.15 with respect to non-coal pollution. The opening of a coal plant, however, triples the elasticity, to more than -0.45, indicating that coal plant emissions are particularly harmful to agricultural productivity. This is consistent with coal emissions being quite different from other forms of pollution; for example, coal emissions contain much more mercury than other forms of pollution.

@@

Figure \@ref(fig:plants) in the appendix displays the location of coal plants in India for two specific years: 1990 and 2010. There is a clear increase in the prevalence of coal plants across the country over the two decades. Additionally, the overall capacity from coal plants increased from `r paste0(reportDecOne(sum(plants1990_capacity$capacity)))` gigawatts to `r paste0(reportDecOne(sum(plants2010_capacity$capacity)))` gigawatts, an increase of `r paste0(reportPercent((sum(plants2010_capacity$capacity) - sum(plants1990_capacity$capacity))/sum(plants1990_capacity$capacity)))` percent in just 20 years. This increase is driven both by the roll out of new plants as well as the construction of new units at existing plants, the latter of which is not shown in Figure \@ref(fig:plants).

@@

[^10]: I define "towards" as within five degrees to help take into account that I am using village centroids, using the x and y components of wind speed to calculate angles.

The top panel in Figure \@ref(fig:windexample) in the appendix is an example wind direction raster. The raster shows wind direction for the entirety of India on January 1st, 2010, as well as the location of coal power plants on that date. The prevailing winds on the date differ across the country and, though not shown in the figure, across days. This means that the overall exposure for a given area to coal plant emissions changes across time. I also pull data on particulate matter to help validate this proxy for exposure. This data comes from @hammer2020global and I also aggregate this to the village using the same method detailed above.

I am primarily interested in the effects of pollution on agricultural productivity. We might consider a regression of the following form:

\begin{gather} y_{it} = \alpha_{i} + \gamma_{t} + \beta PM_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it}, \end{gather}

where $y_{it}$ is outcome $y$ for unit $i$ at time $t$, $\alpha$ is a geographic fixed effects, $\gamma_t$ is time fixed effects, $PM_{it}$ is the concentration of particulate matter 2.5, $rain_{it}$ is total rainfall during the season, and $temp_{it}$ is average temperature during the season.

However, even with the fixed effects, this regression likely suffers from endogeneity; places with more pollution may be very different from places with less pollution. It is difficult to sign the direction of the bias a priori, because it depends on how the coal plants are placed. For example, if coal plants are placed in places in need of power -- for example, due to rapid growth -- then the estimates might be upwards biased. On the other hand, if official purposefully place coal plants in poorer areas for political reasons, then the estimates might be downwards biased.

So what are the characteristics of areas near the construction of new coal plants? Table \@ref(tab:plantresultstable) presents four separate regressions. The first two columns use 1991 census data -- downloaded from SHRUG -- and the second two columns use 2001 census data, all at the village level. The first column presents results from regressing a dummy for whether there is a coal plant within 30km in 1991 on several census variables. Villages with higher populations and higher literacy in 1991 are more likely to be near a coal plant at the time. Interestingly, however, they are less likely to have a paved road.

Given the likely endogeneity, I instead turn to a plausibly exogenous source of variation in pollution: wind direction. While the location of a coal plant might be endogenous, the amount of pollution received by a specific village in a given year can vary randomly based on wind direction. By counting the number of days in a given year that a village is downwind from a polluted area and conditioning on village fixed effects and weather (rainfall and temperature), variation across years is plausibly exogenous.

Consider the reduced form effect of wind exposure on agricultural productivity:
\begin{gather} y_{it} = \alpha_{i} + \gamma_{t} + \beta wind_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it}, \end{gather}

where variables are defined identically to above but with wind as the primary variable of interest instead of pollution. Combining these into a two-stage procedure, we can instrument for pollution using wind direction; small changes in wind direction can lead to large changes in pollution exposure for villages located near point sources of pollution, allowing me to identify the causal effects of pollution by isolating the random component driven by wind direction. This is similar in spirit to @deryugina2019mortality.

It is worth taking the time to validate the use of wind direction as a measure of exposure to pollution. Consider the data used for the regressions presented in Table \@ref(tab:pollutiontable) of the appendix. The outcome variable is particulate matter -- specifically, $\mathrm{PM_{2.5}}$, which is particulate matter no larger than 2.5 micrometers in diameter -- which comes from @hammer2020global. $\mathrm{PM_{2.5}}$ is one of the harmful byproducts from coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.[^12] This data is available at the monthly level, so in Table \@ref(tab:pollutiontable), I aggregate wind exposure to the monthly level, as well.

### Identifying effects of pollution from coal plants

I use the location of coal plants to identify locations of high pollution, but there can be other pollution sources located near coal plants, as well. I nonetheless want to identify the effects of pollution coming specifically from coal plants. I do this by combining the two-stage procedure above with a differences-in-differences framework. Specifically, I create a new variable identifying the exact year the coal plant opens in a given area and calculate the change in the effects of pollution on agricultural productivity.

Consider the following regression of interest:
\begin{gather} 
y_{it} = \alpha_{i} + \gamma_{t} + \omega pm_{it} + \phi open_{it} + \beta pm_{it}\times open_{it} + \delta rain_{it} + \psi temp_{it} + \varepsilon_{it},
\end{gather}

where the interaction term $pm_{it}\times open_{it}$ indicates the change in the effect of pollution following the opening of the coal plant. Note that this is not the same as saying that pollution increased; instead, it is focusing on the change in the effect of a change in pollution of one unit on the outcome of interest, agricultural yield. I instrument for both $pm_{it}$ and $pm_{it}\times open_{it}$ using wind and the interaction between wind and the open indicator, or $wind_{it}\times open_{it}$.

This section presents the main results of this paper. I present two main sets of results: one set looking at agricultural productivity and another looking at labor allocation.

I first present results for agricultural productivity, with the reduced form. The outcome of all the regressions in this section is the log of agricultural land productivity, which is defined as tons per hectare. The first set of results are in Table \@ref(tab:yieldtable). The unit of analysis is the village-year -- from 2002 to 2013 -- focusing only on the monsoon (kharif) season. The data used to construct the yield measure is calculated using five months of data, so the wind variable is defined for the same five months and is the sum of days in which the wind is pointed towards the village centroid from a coal plant.

The first column presents the most simple results, with only the wind exposure variable and the fixed effects. The second column adds a rainfall and average temperature variable, while the third column includes higher-order weather variables, as well. While the coefficient decreases slightly in magnitude as we move from the first to the second column and the second to the third column, the decrease is relatively small. There is a consistent, negative effect of wind exposure on agricultural productivity.

# Table
colnames(yield1reducedform) <- c("(1)", "(2)", "(3)")
kable(
      yield1reducedform, caption = "Wind direction and agricultural productivity",
      align = "ccc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions is estimated agricultural yield, defined as tons per hectare. * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:4),width = "2.5cm") %>%
  row_spec(13, bold = TRUE) %>%
  row_spec(15, hline_after = TRUE) %>%
  #add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
  kable_classic_2()

To put the size of the coefficient in context, it is worth digging a little deeper into the exposure variable. The mean within-village absolute deviation in wind exposure is approximately 2.42 days, meaning that the average change in agricultural productivity from year-to-year due to nothing but changes in wind patterns carrying particulate matter is around 1.5 percent relative to the mean. This implies that swings in agricultural productivity -- due to wind direction -- of several percent are probably quite common, since the absolute deviation includes deviations below and above the mean.

# Table
colnames(yield2naive) <- c("(1)", "(2)", "(3)")
kable(
      yield2naive, caption = "Particulate matter (PM 2.5) and agricultural productivity",
      align = "ccc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions is estimated agricultural yield, defined as tons per hectare. * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:4),width = "2.5cm") %>%
  row_spec(5, bold = TRUE) %>%
  row_spec(7, hline_after = TRUE) %>%
  #add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
  kable_classic_2()

Of course, it is not wind, per se, that we are interested in. Instead, we are actually interested in better understanding the effects of exposure to pollution on agricultural productivity. Table \@ref(tab:yieldtablepm) presents the same regressions as Table \@ref(fig:yieldtable), with a key difference: instead of wind exposure, it uses logged average concentration of $\mathrm{PM_{2.5}}$ -- from @hammer2020global -- during the season as the key explanatory variable. There is again a consistent, negative relationship between particulate matter and agricultural productivity. Since both variables are logged, the coefficients are similar to elasticities, indicating that a one-percent increase in pollution leads to a 0.03-to-0.04 decrease in agricultural productivity.

Recall, however, the results from above: coal plants were placed in places with higher average agricultural productivity. This means that the simple regressions in \@ref(tab:yieldtablepm) might actually be biased upwards: although pollution might negatively affect productivity, pollution is higher in more productive areas due to the endogenous placement of coal plants in the study areas. To recover the actual effect of pollution, we need an instrument. If, conditional on the fixed effects and weather variables, wind direction is exogenous and only affects agricultural productivity through pollution, then it is a valid instrument for pollution.

I present the IV results in Table \@ref(tab:yieldtabletwo), with the first-stage F-statistic in the bottom. The first-stage results make clear how strong of an instrument wind is; all three F-statistics are above 1,000, meeting even the higher bar for the first-stage F-test argued for by @lee2022valid. Also note that the coefficients are quite different from those in Table \@ref(tab:pollutiontable) because the unit of analysis in Table \@ref(tab:yieldtabletwo) is the season, while in Table \@ref(tab:pollutiontable) the unit of analysis is the month.

The second-stage results first indicate that $\mathrm{PM_{2.5}}$ is a very strong predictor of agricultural productivity. Using the results in column three, a one percent increase in particulate matter leads to a 0.73 percent decrease in agricultural productivity. To put this in further perspective, the mean particulate matter concentration is just short of 40 (39.68) in the monsoon sample, while the within-village absolute deviation in particulate matter is 3.98, meaning year-to-year swings of 7-to-14 percent in agricultural productivity due to pollution exposure are common.

# Table
colnames(yield3ivmain) <- c("(1)", "(2)", "(3)")
rownames(yield3ivmain)[8] <- "F (first stage)"
kable(
      yield3ivmain, caption = "Pollution and agricultural productivity, IV estimates",
      align = "ccc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions is estimated agricultural yield, defined as tons per hectare. * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:4),width = "2cm") %>%
  row_spec(5, bold = TRUE) %>%
  row_spec(7, hline_after = TRUE) %>%
  #add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
  kable_classic_2()

The first two columns split the sample based on median exposure. In other words, for each village, I calculate the median exposure (based on wind) and then use this single value for each village to split the sample into villages above and below median exposure. The coefficient in the first column is 2.3 times larger than the coefficient in the second column, indicating much larger effects of pollution in places where it is more common. One possible explanation is that exposure is cumulative, leading to larger and larger effects. I return to this in column five.

Columns three and four split the sample based on median yield at the very beginning of the sample (the first year available, 2002-2003). Column three presents results for villages above the median while column four presents results for villages below the median. The largest effects are apparently concentrated in the upper half of the distribution, meaning the most productive areas are also most susceptible to the effects of pollution.

Given ongoing changes in the climate, a key question is how welfare will change as shocks become more and more common. The last column looks at possible multiplicative effects of shocks by interacting pollution with rainfall and seeing whether the effects compound one another. Since there are now two endogenous regressors, I instrument for both using wind exposure and an interaction between wind exposure and rainfall.

# Table
colnames(yield5heterogeneity) <- c(">p(50)", "<=p(50)", ">p(50)", "<=p(50)", "")
rownames(yield5heterogeneity)[c(1, 3)] <- c("PM 2.5", "PM times rain")
kable(
      yield5heterogeneity, caption = "Heterogeneity in the effects of pollution on productivity",
      align = "ccccc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions is estimated agricultural yield, defined as tons per hectare. The first two columns split the sample based on the median exposure (by village). The second two columns split the sample based on initial -- the first year available -- agricultural productivity. * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "4.5cm") %>%
  column_spec(c(2:6),width = "1.8cm") %>%
  row_spec(c(8), hline_after = TRUE) %>%
  row_spec(c(6), bold = TRUE) %>%
  add_header_above(c(" " = 1, "wind" = 2, "yield" = 2, " " = 1)) %>%
  kable_classic_2()

In the appendix, I present in Table \@ref(tab:yieldtableleads) as a robustness check similar regressions but also including wind/pollution leads. The negative effect is only seen with current pollution, not future pollution.

Table \@ref(tab:labortable) presents results for labor allocation over the last seven days using the National Sample Survey and the methodology described above. Since the labor recall is for seven days, labor allocation ranges from zero to seven, as does the wind variable. Recall that the data is at the district level, not the village level. However, it is also more temporally disaggregated, with rolling interviews throughout the year-long survey. Figure \@ref(fig:laborplot) in the appendix shows this temporal distribution across the five waves of the survey. Since pollution is not disaggregated at a daily level, I instead stick to the reduced form results with wind.

```{r labortable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(labor1all) <- c("all", "all", "self", "wage", "farm", "non-farm")
colnames(labor2rural) <- c("all", "all", "self", "wage", "farm", "non-farm")

mat1 <- matrix(c("", "", "", "", "", ""), nrow = 1)
colnames(mat1) <- c("all", "all", "self", "wage", "farm", "non-farm")
rownames(mat1) <- c("panel a: all")
mat2 <- matrix(c("", "", "", "", "", ""), nrow = 1)
colnames(mat2) <- c("all", "all", "self", "wage", "farm", "non-farm")
rownames(mat2) <- c("panel b: rural")

labortablerural <- rbind(mat1, labor1all[c(1:2, 7:8),], mat2, labor2rural[-c(3:6),], labor2rural[c(3:6),])

The top panel presents results for all respondents during the five-month monsoon season while the bottom panel presents results for respondents in rural areas. Interestingly, we do not see any signs of large, negative effects of wind exposure on labor allocation across any sectors, with the possible exception of wage employment in rural areas. However, the coefficient for self-employment in rural areas is of the same magnitude and *positive*, but imprecisely estimated, making it hard to come to any definitive conclusions, at least in the aggregate.

While overall results for labor allocation are inconclusive, perhaps we can learn more by looking at temporally disaggregated results, focusing on May through December. Although the yield data was created by looking at five specific months of the year -- June through October -- this is partially because this is the time of year when remote-sensing data is most able to pick up overall changes in crop cover. Preparation, however, starts before that, with operations such as ploughing often taking place prior to June.

Since the NSS data has survey dates and the both the pollution and wind data can be aggregated to the month, we can look at how exposure during different months of the agricultural season (differentially) affect labor allocation and overall agricultural productivity. I begin with labor allocation in Table \@ref(tab:labor3month). As before, the exposure measure is just the wind variable (over the last seven days), since I can directly match daily wind direction with the timing of the survey, but I cannot do that for pollution, which is monthly. The first column presents results for farm labor and the second column presents results for non-farm labor.

Higher exposure leads to significantly less farm labor labor in June and July, but there are no discernible effects in other months of the year. There are likewise no discernible effects on non-farm labor, but some of the coefficients are rather imprecisely estimate. To put the effects for June and July in context and taking the coefficients at face value, an additional day of wind in the last seven days decreases farm labor in June by around `r paste0(reportPercent(0.136/1.209))` percent of the mean and in July by around `r paste0(reportPercent(0.285/1.671))` percent of the mean.

```{r labor3month, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(labor3month) <- c("farm", "non-farm")
kable(
      labor3month, caption = "Wind and labor allocation, by month",
      align = "cc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the village level. The outcome in column one is .  * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:3),width = "2cm") %>%
  row_spec(c(18), hline_after = TRUE) %>%
  row_spec(c(16), bold = TRUE) %>%
  kable_classic_2()

I perform a similar exercise with yield, the results of which are in Table \@ref(tab:yield4monthly). I include only the five months of the year that are used in the calculation of the yield estimates, or June through October. The results are again from two-stage least squares, with the F-statistics for each month presented at the bottom; monthly wind remains a strong predictor of pollution. Column one presents results without controls for weather while column two presents results with controls for weather; the differences in coefficients make clear how important the inclusion of weather is here.

Focusing on column two, there is only one month where we see a clear effect of pollution on yield: June. Given that we also see negative effects of pollution on farm labor allocation in June, this is suggestive -- but not conclusive -- evidence that pollution works through labor. Note that this does not rule out pollution also having direct effects on agricultural productivity, however.

```{r yield4monthly, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(yield4monthly) <- c("(1)", "(2)")
kable(
      yield4monthly, caption = "Pollution and productivity, by month",
      align = "cc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions is estimated agricultural yield, defined as tons per hectare. * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:3),width = "2cm") %>%
  row_spec(12, bold = TRUE) %>%
  row_spec(14, hline_after = TRUE) %>%
  #add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
  kable_classic_2()

A final caveat about these month-specific results is that they do not take into account possible interactions between different months. For example, yield might respond strongly to average pollution in June and July, which would not be captured in Table \@ref(tab:yield4monthly).

## Coal plants or overall pollution?
